---
title: "Plots of Sunspots: A Project in Fulfillment of Course 9 of the Data Science Specialty"
author: "Justin G."
date: "September 16, 2017"
output:
  ioslides_presentation: default
  slidy_presentation: default
---

```{r initialize, echo=FALSE,  warning=F, message=F}
library(knitr); library(rmarkdown)
```

## Rationale
- Many look at sunspot cycles to study space weather and astronomy.
- Time series of sunspots reveal cyclical patterns.
- Mean relative sunspot numbers from 1749-1983 were collected at the Swiss Federal Observatory until 1960,
and then by the Tokyo Astronomical Observatory.
- These data are freely available in R by typing *sunspots*.  More dataset information: type *?sunspots* in R.

## This Project
- Hosts interactive plot on [http://shinyapps.io](http://shinyapps.io), allowing the user to view a plot of mean sunspot numbers by month and year of his/her choice.
- Uses sliders to modify plots (not shown in this presentation).
- Can be used as an **initial** step for subsequent data analysis, correlation of sunspots with other environmental phenomena, etc.
- Y-axis values change in response to user's input for X axis (month or yr, depending on plot)
- Of course, it is up to the user to choose a month or yr (see actual app).

## Plot examples - choose a year
```{r one_plot, echo=TRUE, warning=F, message=F, fig.cap="1st plot: choose a year (here: 1751)"}
sunspotsbymonth<-matrix(sunspots, ncol=13)
plot(sunspotsbymonth[3,], xlim=c(1,12), xlab="Calendar Month", 
         ylab="Number of sunspots per month", pch=19, cex=2, col="red")
```

## Plot examples - choose a month
```{r two_plots, echo=TRUE, warning=F, message=F, fig.cap="2nd plot: choose a month (here July)"}
sunspotsbymonth<-matrix(sunspots, ncol=13)
plot(sunspotsbymonth[,7], xlab="Calendar Year", xaxt="n",
        ylab="Number of sunspots per year", pch=17, cex=1.5, col="blue")
    axis(1, at=1:235, labels=c(1749:1983))
```

## Conclusions
- **Interactive** Shiny app available at: https://justgo129.shinyapps.io/course9project/
- Looking forward to your view!

